canaper: Categorical analysis of neo- and paleo-endemism in R

### 2022-09-30 ms/ms.Qmd
Update MS

Joel H. Nitta^1,\*^, Shawn W. Laffan^2^, Brent D. Mishler^3^, and Wataru Iwasaki^1,4,5,6,7^

@@

^5^Atmosphere and Ocean Research Institute, The University of Tokyo, Chiba, Japan

^6^Institute for Quantitative Biosciences, The University of Tokyo, Tokyo, Japan

^7^Collaborative Research Institute for Innovative Microbiology, The University of Tokyo, Tokyo, Japan

@@

@@

data(phylocom)

# The original data includes abundance
phylocom$comm

@@

```{r}
#| label: acacia-stats
#| include: false
data(acacia)

@@

```{r}
#| label: cpr-iter-sim-demo
data(acacia)

@@

```{r}
#| label: cpr-rand-test-demo
rand_test_results <- cpr_rand_test(
  phylocom$comm, phylocom$phy, null_model = "curveball")

@@

```{r}
#| label: cpr-rand-comm-parallel-demo
#| eval: false
library(future)

@@

a <- ggplot(acacia_canape, aes(x = long, y = lat, fill = endem_type)) +
  geom_tile() +
  # use same colors as Misher et al. 2014
  # (for CVD-safe colors see cpr_endem_cols or cpr_endem_cols_2)
  scale_fill_manual(values = mishler_endem_cols) +
  guides(
    fill = guide_legend(title.position = "top", label.position = "bottom")
  ) +
  coord_fixed() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.minor = element_blank()
  )

@@

# Tally number of significant cells in canper vs. Biodiverse
sig_tally <-
  comp_table %>%
  mutate(
    endem_type = str_replace_all(
      endem_type, "not significant", "not_significant") %>%
      as.factor() %>%
      fct_collapse(significant = c("neo", "paleo", "mixed", "super"))
      ) %>%
  rename(canaper = n_canaper, biodiverse = n_biodiverse) %>%
  pivot_longer(names_to = "analysis", values_to = "n_sites", -endem_type) %>%
  group_by(endem_type, analysis) %>%
  summarize(n_sites = sum(n_sites), .groups = "drop") %>%
  split(.$analysis) %>%
  purrr::map(., ~split(., .$endem_type))
```

@@

This research supported in part by Japan Society for the Promotion of Science (Kakenhi) Grant numbers 16H06279 and 22H04925.
Members of the Iwasaki Lab (The University of Tokyo) provided helpful comments while the package was in development.
Luis Osorio and Klaus Schliep reviewed the code as part of submission to ROpenSci.

@@

Joel H. Nitta conceived the ideas and designed methodology; Joel H. Nitta and Shawn W. Laffan analyzed the data; Joel H. Nitta led the writing of the manuscript.
All authors contributed critically to the drafts and gave final approval for publication.

>>>>

Joel H. Nitta^1,\*^, Shawn W. Laffan^2^, Brent D. Mishler^3^, and Wataru Iwasaki^1,4,5,6,7,8^

@@

^5^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, Tokyo, Japan

^6^Atmosphere and Ocean Research Institute, The University of Tokyo, Chiba, Japan

^7^Institute for Quantitative Biosciences, The University of Tokyo, Tokyo, Japan

^8^Collaborative Research Institute for Innovative Microbiology, The University of Tokyo, Tokyo, Japan

@@

# Installation

The stable version of `canaper` can be installed from CRAN, and the latest development version from r-universe (<https://r-universe.dev/>).

```{r}
#| label: installation
#| eval: false

# Stable version
install.packages("canaper")

# Development version
install.packages("canaper", repos = "https://ropensci.r-universe.dev")
```

@@

# The original data includes abundance
data(phylocom)
phylocom$comm

@@

```{r}
#| label: acacia-stats
#| include: false
set.seed(12345)

data(acacia)

@@

```{r}
#| label: cpr-iter-sim-demo
set.seed(12345)

data(acacia)

@@

```{r}
#| label: cpr-rand-test-demo
set.seed(12345)

rand_test_results <- cpr_rand_test(
  phylocom$comm, phylocom$phy, null_model = "curveball")

@@

```{r}
#| label: cpr-rand-comm-parallel-demo
#| eval: false
library(future)
set.seed(12345)

@@

a <- ggplot(acacia_canape, aes(x = long, y = lat, fill = endem_type)) +
  geom_tile() +
  # Use same colors as Misher et al. 2014
  # (for CVD-safe colors see cpr_endem_cols)
  scale_fill_manual(values = mishler_endem_cols) +
  guides(
    fill = guide_legend(title.position = "top", label.position = "bottom")
  ) +
  coord_fixed() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.minor = element_blank()
  )

@@

# Tally number of significant cells in canaper vs. Biodiverse
sig_tally <-
  comp_table %>%
  mutate(
    endem_type = str_replace_all(
      endem_type, "not significant", "not_significant") %>%
      as.factor() %>%
      fct_collapse(significant = c("neo", "paleo", "mixed", "super"))
      ) %>%
  rename(canaper = n_canaper, biodiverse = n_biodiverse) %>%
  pivot_longer(names_to = "analysis", values_to = "n_sites", -endem_type) %>%
  group_by(endem_type, analysis) %>%
  summarize(n_sites = sum(n_sites), .groups = "drop") %>%
  split(.$analysis) %>%
  purrr::map(., ~split(., .$endem_type))
```

@@

This research supported in part by Japan Society for the Promotion of Science (Kakenhi) Grant numbers 16H06279, 22H04925, and 22K15171.
Members of the Iwasaki Lab (The University of Tokyo) provided helpful comments while the package was in development.
Luis Osorio and Klaus Schliep reviewed the code as part of submission to ROpenSci.

@@

Joel H. Nitta conceived the ideas and designed methodology; Joel H. Nitta and Shawn W. Laffan analyzed the data; Joel H. Nitta led the writing of the manuscript; Brent D. Mishler edited the manuscript; Wataru Iwasaki provided resources and supervised the study.
All authors contributed critically to the drafts and gave final approval for publication.
